---
title: "Example analysis using the `fsca` package"
output: rmarkdown::html_vignette
bibliography: "PhD.bib"
csl: "apa-6th-edition.csl"
vignette: >
  %\VignetteIndexEntry{example-analysis}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```


This vignette illustrates a complete analysis of syntactic complexity in one text 
from start to finish. 

# Parse Text

The text file `syntaxe-wiki.txt` contains some text from the [French wikipedia
page about Syntax](https://fr.wikipedia.org/wiki/Syntaxe).

> La syntaxe est, à l'origine, la branche de la linguistique qui étudie la façon dont les mots se combinent pour former des phrases ou des énoncés dans une langue. On distingue la syntaxe, qui concerne les expressions [les mots], de la sémantique, qui concerne ce qui est visé par les expressions [le sens, la signification/les choses]. Le terme de syntaxe est aussi utilisé en informatique, où sa définition est similaire, modulo une terminologie différente. Ainsi, la syntaxe est le respect, ou le non-respect, de la grammaire formelle d'un langage, c'est-à-dire des règles d'agencement des lexèmes (en informatique, ce sont des entités lexicales d'un langage informatique) en des termes plus complexes, souvent des programmes. Dans la théorie des langages formels, ce qui joue le rôle de lexème est en général appelé lettre ou symbole, et les termes produits sont appelés mots.

This text file was POS-tagged with MElt Tagger [@Denis2012] and dependency 
parsed with Malt Parser [@Nivre2006]. See [here](http://alpage.inria.fr/statgram/frdep/fr_stat_dep_malt.html) 
for instructions. 

# Extract parsed text

The output of the pre-processing chain is a text file containing the sentences in CONLL format. 
Each sentence is separated by an empty line.

```{r}
head(read.table("syntaxe-wiki.txt.outmalt"))
```

This output can be reformatted with the `getParse()` function. This function separates each of the 
sentences into a list and prepares the data frames for analysis (changing column names, data type etc.).

```{r}
library(fsca)
list <- getParse("syntaxe-wiki.txt.outmalt")
head(list[[1]])
```

# Analyze text

Syntactic units can then be extracted using the `getUnits()` function on individual sentences 
(i.e. by subsetting the list of sentences):

```{r}
getUnits(list[[1]], paste.tokens = TRUE)
```

Syntactic complexity measures can be directly calculated for the entire list 
using the `getMeasures()` function:

```{r}
getMeasures(list)
```


# References
